---
title: "ITT Liner Mixed Effect Model"
author: "Catherine"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
options(servr.daemon = TRUE)
knitr::opts_chunk$set(dpi= 300, fig.align = 'center')
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer); library(tidyverse)
library(lme4);library(sjPlot);library(gtsummary);
library(ggpubr)
load("join_yes_perm1.rdata")
style_mono_accent(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "400", "400"),
  code_font_google   = google_font("Fira Mono"),
  base_font_size = "16px"
)
```

### Descriptive for matched data (N = 1149, bootstrap once)
.pull-left[
```{r echo=F}
#tab_xtab(df_plot$project, df_plot$visit)

df_match %>% 
  select(-doa_bl, -subject_code, -hub, -k6_tier_sc) %>% 
  mutate(tier = ifelse(tier %in% c("T0", "T1"), "T1", tier)) %>% 
  tbl_strata(strata = tier, 
             .tbl_fun = ~ .x %>% 
               gtsummary::tbl_summary(by = project,
                                      statistic = all_continuous() ~ "{mean} ({sd})"))
```
]
.pull-right[
```{r echo=F}
df_match %>% 
  select(-doa_bl, -subject_code, -hub, -k6_tier_sc) %>% 
  mutate(tier = ifelse(tier %in% c("T0", "T1"), "T1", tier)) %>% 
  gtsummary::tbl_summary(by = project,statistic = all_continuous() ~ "{mean} ({sd})")

```

]

```{r echo=F, warning=F, include=F}
df_plot$project <- relevel(df_plot$project, ref = "Control")

lme1 <- lmer(values ~ project*visit + (1|doa_bl) + (1|prof_service), 
             data = df_plot)
lme2 <- lmer(values ~ project*visit + (1|doa_bl) + (1|prof_service), data = df_plot %>% filter(tier %in% c("T0", "T1")))
lme3 <- lmer(values ~ project*visit + (1|doa_bl) + (1|prof_service), data = df_plot %>% filter(tier %in% c("T2")))
lme4 <- lmer(values ~ project*visit + (1|doa_bl) + (1|prof_service), data = df_plot %>% filter(tier %in% c("T3")))
```

---
.pull-left[
### All tiers
```{r echo=F}
plot_model(lme1, type = "pred", terms = c("visit", "project")) +
  geom_line() +
  ggtitle("Predicted value of K6") +
  geom_text(aes(label = round(predicted, 3)), 
            position = position_dodge(width = 1),
            vjust = 1) +
  theme(legend.position = "bottom") + 
  theme_xaringan()
```

]
.pull-right[
```{r echo=F}
tab_model(lme1)
```
]
---
.pull-left[
### T1
```{r echo=F}
plot <- plot_model(lme2, type = "pred", terms = c("visit", "project"))
plot_model(lme2, type = "pred", terms = c("visit", "project")) +
  geom_line() +
  ggtitle("Predicted value of K6") +
  geom_text(aes(label = round(predicted, 3)), 
            position = position_dodge(width = 1),
            vjust = 1) +
  theme(legend.position = "bottom") + 
  theme_xaringan()
```

]
.pull-right[
```{r echo=F}
tab_model(lme2)
```
]

---
.pull-left[
### T2
```{r echo=F}
plot_model(lme3, type = "pred", terms = c("visit", "project")) +
  geom_line() +
  ggtitle("Predicted value of K6") +
  geom_text(aes(label = round(predicted, 3)), 
            position = position_dodge(width = 1),
            vjust = 1) +
  theme(legend.position = "bottom") + 
  theme_xaringan()
```

]
.pull-right[
```{r echo=F}
tab_model(lme3)
```
]

---
.pull-left[
### T3
```{r echo=F}
plot_model(lme4, type = "pred", terms = c("visit", "project")) +
  geom_line() +
  ggtitle("Predicted value of K6") +
  geom_text(aes(label = round(predicted, 3)), 
            position = position_dodge(width = 1),
            vjust = 1) +
  theme(legend.position = "bottom") + 
  theme_xaringan()
```
]
.pull-right[
```{r echo=F}
tab_model(lme4)
```
]
---
### Descriptive for crude data (YES = 1551, Hub = 1225)
#### Model built using different setting
- Include all age in hub
- Add baseline K6 as a random intercept
- Bootstrapped parameters (100 times)

.pull-left[
```{r echo=F}
df_join_pre <- 
      df_2m_last %>% 
      janitor::clean_names() %>% 
      select(subject_code, k6_last) %>% 
      left_join(df_2m_first %>% 
                  janitor::clean_names() %>% 
                  select(subject_code, k6_first, k6_tier_sc, 
                         doa_bl, prof_service),
                by = "subject_code") %>% 
      mutate(project = 1) %>% 
      union_all(
        df_yes_3m %>% 
          janitor::clean_names() %>% 
          select(subject_code, k6_last) %>% 
          left_join(df_yes_bl %>% 
                      janitor::clean_names() %>% 
                      select(subject_code, k6_first, k6_tier_sc,
                             doa_bl, prof_service),
                    by = "subject_code") %>% 
          mutate(project = 0) 
      ) %>% 
      ungroup()
    df_join <- 
      df_join_pre %>%
      mutate(tier = case_when(k6_tier_sc >= 0 & k6_tier_sc < 9 ~ "T0",
                              k6_tier_sc >= 9 & k6_tier_sc < 11 ~ "T1",
                              k6_tier_sc >= 11 & k6_tier_sc < 15 ~ "T2",
                              k6_tier_sc >= 15 & k6_tier_sc <= 24 ~ "T3"),
             hub = gsub("[^a-zA-Z]", "", subject_code),
             project = if_else(project == 1, "Hub","Control") %>% 
               factor(., levels = c("Hub","Control"))) %>% 
      filter(complete.cases(.)) %>% 
      arrange(project, desc(k6_tier_sc))
df_join_flip <- df_join %>% pivot_longer(k6_first:k6_last,
                     names_to = "visit", values_to = "values") %>% 
  mutate(visit = factor(visit, levels = c("k6_first", "k6_last")))

df_join_flip$project <- relevel(df_join_flip$project, ref = "Control")

#tab_xtab(df_join_flip$project, df_join_flip$visit)
df_join %>% 
  select(-doa_bl, -subject_code, -hub, -k6_tier_sc) %>% 
  select(k6_first, k6_last, everything()) %>% 
  mutate(tier = ifelse(tier %in% c("T0", "T1"), "T1", tier)) %>% 
  tbl_strata(strata = tier, 
             .tbl_fun = ~ .x %>% 
               gtsummary::tbl_summary(by = project,
                                      statistic = all_continuous() 
                                      ~ "{mean} ({sd})"))
```
]
.pull-right[
```{r echo=F}
df_join %>% 
  select(-doa_bl, -subject_code, -hub, -k6_tier_sc) %>% 
  select(k6_first, k6_last, everything()) %>% 
  gtsummary::tbl_summary(by = project, 
                         statistic = all_continuous() ~ "{mean} ({sd})")
```
]
```{r echo=F, warning=F, include=F}
lme1 <- lmer(
  values ~ project*visit  + (1|doa_bl) + (1|prof_service) + (1|k6_tier_sc), 
             data = df_join_flip)

lme2 <- lmer(
  values ~ project*visit + (1|doa_bl) + (1|prof_service) + (1|k6_tier_sc), 
             data = df_join_flip %>% filter(tier %in% c("T0", "T1")))

lme3 <- lmer(
  values ~ project*visit + (1|doa_bl) + (1|prof_service) + (1|k6_tier_sc), 
             data = df_join_flip %>% filter(tier %in% c("T2")))

lme4 <- lmer(
  values ~ project*visit + (1|doa_bl) + (1|prof_service) + 
    (1|k6_tier_sc), 
             data = df_join_flip %>% filter(tier %in% c("T3")))
```
---
### All tiers
.pull-left[
```{r echo=F, warning=F}
set.seed(1234)
boot <- 
  parameters::bootstrap_parameters(lme1, iterations = 100, bootstrap = T, test = "p")
null.model <- lmer(
  values ~ 1 + (1|doa_bl) + (1|prof_service) + (1|k6_tier_sc), 
             data = df_join_flip)
r2 <- 
  (summary(null.model)[["sigma"]]^2 - summary(lme1)[["sigma"]]^2)/
  summary(null.model)[["sigma"]]^2
f2 <- round(r2/(1-r2), 3)
# A practical guide to calculating Cohen’s f2, a measure of local effect size, from PROC MIXED
boot.df <- tibble(visit = rep(c("k6_first", "k6_last"), 2),
                  project = rep(c("Hub", "Control"), each = 2),
                  values = c(
                    boot$Coefficient[1] + boot$Coefficient[2],
                    boot$Coefficient[1] + boot$Coefficient[2] + 
                    boot$Coefficient[3] + boot$Coefficient[4],
                    boot$Coefficient[1],
                    boot$Coefficient[1] + boot$Coefficient[3]))

ggplot(boot.df,
       aes(x = visit, y = values, color = project, group = project)) +
  geom_point() +
  geom_line() +
  scale_y_continuous(limits = c(0, 24)) +
  geom_text(aes(label = round(values, 3)), vjust = 0.75,
            position = position_dodge(width = 1), size = 5) +
  labs(subtitle = bquote(paste("Cohen's", " ", f[Fixed]^2, " = ", .(f2)))) + 
  theme(legend.position = "bottom") + 
  ggtitle("Predicted Value") +
  theme_xaringan() 
```
]
.pull-right[
```{r echo=F}
boot %>% knitr::kable(caption = "100 Bootstrap Parameters")
```
]

---
### T1
.pull-left[
```{r echo=F, warning=F}
set.seed(1234)
boot <- 
  parameters::bootstrap_parameters(lme2, iterations = 100, bootstrap = T, test = "p")
null.model <- lmer(
  values ~ 1 + (1|doa_bl) + (1|prof_service) + (1|k6_tier_sc), 
             data = df_join_flip %>% filter(tier %in% c("T0", "T1")))
r2 <- 
  (summary(null.model)[["sigma"]]^2 - summary(lme2)[["sigma"]]^2)/
  summary(null.model)[["sigma"]]^2
f2 <- round(r2/(1-r2), 3)
boot.df <- tibble(visit = rep(c("k6_first", "k6_last"), 2),
                  project = rep(c("Hub", "Control"), each = 2),
                  values = c(
                    boot$Coefficient[1] + boot$Coefficient[2],
                    boot$Coefficient[1] + boot$Coefficient[2] + 
                      boot$Coefficient[3] + boot$Coefficient[4],
                    boot$Coefficient[1],
                    boot$Coefficient[1] + boot$Coefficient[3]))
ggplot(boot.df,
       aes(x = visit, y = values, color = project, group = project)) +
  geom_point() +
  geom_line() +
  scale_y_continuous(limits = c(0, 25)) +
  geom_text(aes(label = round(values, 3)), vjust = 0.75,
            position = position_dodge(width = 1), size = 5) +
  labs(subtitle = bquote(paste("Cohen's", " ", f[Fixed]^2, " = ", .(f2)))) +
  ggtitle("Predicted Value") +
  theme(legend.position = "bottom") + 
  theme_xaringan()
```
]
.pull-right[
```{r echo=F}
boot %>% knitr::kable(caption = "100 Bootstrap Parameters")
```
]


---
### T2
.pull-left[
```{r echo=F, warning=F, message=F}
set.seed(1234)
boot <- 
  parameters::bootstrap_parameters(lme3, iterations = 100, bootstrap = T, test = "p")
null.model <- lmer(
  values ~ 1 + (1|doa_bl) + (1|prof_service) + (1|k6_tier_sc), 
             data = df_join_flip %>% filter(tier %in% c("T2")))
r2 <- 
  (summary(null.model)[["sigma"]]^2 - summary(lme3)[["sigma"]]^2)/
  summary(null.model)[["sigma"]]^2
f2 <- round(r2/(1-r2), 3)

boot.df <- tibble(visit = rep(c("k6_first", "k6_last"), 2),
                  project = rep(c("Hub", "Control"), each = 2),
                  values = c(
                    boot$Coefficient[1] + boot$Coefficient[2],
                    boot$Coefficient[1] + boot$Coefficient[2] + 
                      boot$Coefficient[3] + boot$Coefficient[4],
                    boot$Coefficient[1],
                    boot$Coefficient[1] + boot$Coefficient[3]))
ggplot(boot.df,
       aes(x = visit, y = values, color = project, group = project)) +
  geom_point() +
  geom_line() +
  scale_y_continuous(limits = c(0, 25)) +
  geom_text(aes(label = round(values, 3)), vjust = 0.75,
            position = position_dodge(width = 1), size = 5) +
  labs(subtitle = bquote(paste("Cohen's", " ", f[Fixed]^2, " = ", .(f2)))) +
  ggtitle("Predicted Value") +
  theme(legend.position = "bottom") + 
  theme_xaringan()
```
]
.pull-right[
```{r echo=F}
boot %>% knitr::kable(caption = "100 Bootstrap Parameters")
```
]
---
### T3
.pull-left[
```{r echo=F, warning=F}
set.seed(1234)
boot <- 
  parameters::bootstrap_parameters(lme4, iterations = 100, bootstrap = T, test = "p")
null.model <- lmer(
  values ~ 1 + (1|doa_bl) + (1|prof_service) + (1|k6_tier_sc), 
             data = df_join_flip %>% filter(tier %in% c("T3")))
r2 <- 
  (summary(null.model)[["sigma"]]^2 - summary(lme4)[["sigma"]]^2)/
  summary(null.model)[["sigma"]]^2
f2 <- round(r2/(1-r2), 3)

boot.df <- tibble(visit = rep(c("k6_first", "k6_last"), 2),
                  project = rep(c("Hub", "Control"), each = 2),
                  values = c(
                    boot$Coefficient[1] + boot$Coefficient[2],
                    boot$Coefficient[1] + boot$Coefficient[2] + 
                      boot$Coefficient[3] + boot$Coefficient[4],
                    boot$Coefficient[1],
                    boot$Coefficient[1] + boot$Coefficient[3]))
ggplot(boot.df,
       aes(x = visit, y = values, color = project, group = project)) +
  geom_point() +
  geom_line() +
  scale_y_continuous(limits = c(0, 25)) +
  geom_text(aes(label = round(values, 3)), vjust = 0.75,
            position = position_dodge(width = 1), size = 5) +
  labs(subtitle = bquote(paste("Cohen's", " ", f[Fixed]^2, " = ", .(f2)))) +
  ggtitle("Predicted Value") +
  theme(legend.position = "bottom") + 
  theme_xaringan()
```
]
.pull-right[
```{r echo=F}
boot %>% knitr::kable(caption = "100 Bootstrap Parameters")
```
]
